---
title: "Chapter 5 (Exercises)"
author: "Aditya Dahiya"
date: 2021-02-14
subtitle: "Resampling Methods"
execute: 
  echo: true
  warning: false
  error: false
  cache: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
---

# 5.4 Exercises

# Conceptual

## Question 1

For this question, we need to find the value of $\alpha$ which minimizes
the expression:\
$$
\begin{aligned}
Total \ Variance &= Var(\alpha X + (1-\alpha) Y) \\
&= Var(\alpha X) + Var((1-\alpha)Y) + 2.Cov(\alpha X, (1-\alpha)Y) \\
&= \alpha^2Var(X) + (1-\alpha)^2Var(Y) + 2Cov(X,Y)(\alpha(1-\alpha)) \\
&= \alpha^2 \sigma_X^2 + (1-\alpha)^2 \sigma_Y^2 + 2 \sigma_{XY}(\alpha(1-\alpha)) \\
\end{aligned}
$$

Now, to find the value of $\alpha$ which minimizes this total variance,
we take it's first partial derivative with respect to $\alpha$ and set
it equal to zero:--

$$
\begin{aligned}0 &= \frac{\partial}{\partial \alpha} \Big( \alpha^2 \sigma_X^2 + (1-\alpha)^2 \sigma_Y^2 + 2 \sigma_{XY}(\alpha(1-\alpha)) \Big) \\0 &= 2\alpha \sigma_X^2 + 2(1-\alpha) \frac{\partial (1-\alpha)}{\partial \alpha} \sigma_Y^2 + 2 \sigma_{XY} \Big( \frac{\partial (1-\alpha)}{\partial \alpha}\alpha + \frac{\partial (\alpha)}{\partial \alpha}(1-\alpha) \Big)\\0 &= 2\alpha \sigma_X^2 - 2(1-\alpha) \sigma_Y^2 + 2 \sigma_{XY}  (-\alpha + 1 - \alpha) \\0 &= 2\alpha \sigma_X^2 - 2\sigma_Y^2 + 2\alpha \sigma_Y^2 + 2\sigma_{XY}  - 4\alpha \sigma_{XY} \\2 \Big(\sigma_Y^2 - \sigma_{XY} \Big) &= 2 \Big(\sigma_X^2 + \sigma_Y^2  -2 \sigma_{XY}\Big) \alpha \\\alpha &= \frac{\sigma_Y^2 - \sigma_{XY}}{\sigma_X^2 + \sigma_Y^2  -2 \sigma_{XY}} \qquad (5.6)\end{aligned}
$$

Further, to verify that the first partial derivative is indeed the
minimum value of Variance, we find the second partial derivative with
respect to $\alpha$ and see whether it is zero or not.

$$
\begin{aligned}
&= \frac{\partial^2}{\partial^2 \alpha} \Big( \alpha^2 \sigma_X^2 + (1-\alpha)^2 \sigma_Y^2 + 2 \sigma_{XY}(\alpha(1-\alpha)) \Big) \\
&= \frac{\partial}{\partial \alpha} \Big(2\alpha \sigma_X^2 - 2\sigma_Y^2 + 2\alpha \sigma_Y^2 + 2\sigma_{XY}  - 4\alpha \sigma_{XY} \Big) \\
&= 2 \sigma_X^2 -0 + 2 \sigma_Y^2 + 0  - 4 \sigma_{XY} \\
&= 2 \Big( \sigma_X^2 + 2 \sigma_Y^2 - 2 \sigma_{XY} \Big) \\
&= 2 \Big( Var(X-Y) \Big)
\end{aligned}
$$

Now, the variance of any variable is always positive. Hence, the
expression $2Var(X-Y)$ will always be positive. Thus, we can confirm
that the minimum value is at the first partial derivative. Equation 5.6
is thus proved.

## Question 2

**We will now derive the probability that a given observation is part of
a bootstrap sample. Suppose that we obtain a bootstrap sample from a set
of n observations.**

### (a)

**What is the probability that the first bootstrap observation is not
the jth observation from the original sample? Justify your answer.**

The bootstrap method samples each bootstrap observation from the
original sample randomly with replacement. Thus, for each bootstrap
observation, every value in the original sample has an equal probability
of being selected. This probability is simply $^1/_n$. So, the
probability that first bootstrap observation is the $j^{th}$ observation
from the original sample is $^1/_n$. Finally, the probability that first
bootstrap observation is not the $j^{th}$ observation from original
sample is $1 - (^1/_n)$.

### (b)

**What is the probability that the second bootstrap observation is not
the jth observation from the original sample?**

As answered earlier, the bootstrap method samples each bootstrap
observation from the original sample randomly with replacement. Thus,
for each and every bootstrap observation, every value in the original
sample has an equal probability of being selected. This probability is
simply $^1/_n$.\
So, the probability that second (or, in fact any $k^{th}$) bootstrap
observation is the $j^{th}$ observation from the original sample is
$^1/_n$. Finally, the probability that the second bootstrap observation
is not the $j^{th}$ observation from original sample is $1 - (^1/_n)$.

### (c)

**Argue that the probability that the jth observation is not in the
bootstrap sample is** $(1 − 1/n)^n$.

Since the bootstrap method samples each observation as a random sample
(with replacement) from the original sample, it can be said that each
bootstrap sample is an independent random variable. Now, the probability
of joint occurrence of independent events (in this case, not selecting
the $j^{th}$ observation of original sample) $n$ different times is: $$
\begin{aligned}
Probability &= Pr_1 \times Pr_2 \times Pr_3 \times \ ... \times \ Pr_n \\
&= (1 - \ ^1/_n) \ \times \ (1 - \ ^1/_n) \ \times \  (1 - \ ^1/_n) \ \times \ ... \ (1 - \ ^1/_n) \\
&= (1 - \ ^1/_n)^n
\end{aligned}
$$

### (d)

**When n = 5, what is the probability that the j^th^ observation is in
the bootstrap sample?**

The probability of the $j^{th}$ observation being in the bootstrap
sample is (1 - probability of $j^{th}$ observation being not in the
boostrap sample). For $n=5$, this is 0.672. $$
\begin{aligned}
Probability &= 1- (1 - \ ^1/_n)^n \\
&= 1 - (1-0.2)^5 \\
&= 1 - 0.8^5 \\
&= 0.672
\end{aligned}
$$

### (e)

**When n = 100, what is the probability that the jth observation is in
the bootstrap sample?**

The probability of the $j^{th}$ observation being in the bootstrap
sample is (1 - probability of $j^{th}$ observation being not in the
bootstrap sample). For $n=100$, this is 0.634. $$
\begin{aligned}
Probability &= 1- (1 - \ ^1/_n)^n \\
&= 1 - (1-0.01)^{100} \\
&= 1 - 0.99^{100} \\
&= 0.634
\end{aligned}
$$

### (f)

**When n = 10, 000, what is the probability that the j^th^ observation
is in the bootstrap sample?**

As calculated above, the probability of the $j^{th}$ observation being
in the bootstrap sample is (1 - probability of $j^{th}$ observation
being not in the bootstrap sample). For $n=10000$, this is 0.632. $$
\begin{aligned}
Probability &= 1- (1 - \ ^1/_n)^n \\
&= 1 - (1-0.0001)^{10000} \\
&= 1 - 0.9999^{10000} \\
&= 0.632
\end{aligned}
$$

### (g)

**Create a plot that displays, for each integer value of n from 1 to
100, 000, the probability that the j^th^ observation is in the bootstrap
sample. Comment on what you observe.**

We can use the formula derived above to create such a plot in R. The
code and output is shown below. We observe that as $n \to \infty$, the
value of this probability approaches an asymptote of approximately
0.632.

```{r}
n <- 1:100000
Pr <- 1 - (1 - 1 / n)^n
plot(n, Pr)
library(ggplot2)
ggplot() +
  geom_point(aes(n, Pr), shape = 20) +
  labs(
    x = "Value of n",
    y = "Probability (jth observation in bootstrap sample)"
  ) +
  theme_classic() +
  scale_y_continuous(
    breaks = c(0.6, 0.632, 0.7, 0.8, 0.9, 1),
    limits = c(0.6, 1)
  )
```

### (h)

**We will now investigate numerically the probability that a bootstrap
sample of size n = 100 contains the j^th^ observation. Here j = 4. We
repeatedly create bootstrap samples, and each time we record whether or
not the fourth observation is contained in the bootstrap sample. Comment
on the results obtained.**

The R-code and results are displayed below. The Code contains an
implementation of a bootstrap method, where we take 10,000 samples (each
containing 100 numbers) from the list of first 100 positive integers
1,2,3,...,100 with replacement. We then check whether the sample
contains the integer 4. We then find whether each sample contains digit
4, and store the result (TRUE or FALSE) in the $i^{th}$ value of `store`
vector. The mean of `store` then gives the probability that a bootstrap
of size n=100 contains the 4th observation.

```{r, collapse=TRUE}
store <- rep(NA, 10000)
for (i in 1:10000) {
  store[i] <- sum(sample(1:100, replace = TRUE) == 4) > 0
}
mean(store)
```

The results reconcile with out calculated probability. It is very close
to 0.632. Further, a known formula from Calculus states that:
$$\lim_{n \rightarrow \infty}(1 + \ ^x/_n \ )^n = e^x$$ Here, $x=-1$,
and thus the expression is equivalent to $e^{-1}$, which is 0.368. So,
$1 - e^{-1}$ is 0.632 which is the calculated answer for previous two
questions.

## Question 3

**We now review k-fold cross-validation.**

### (a)

**Explain how k-fold cross-validation is implemented.**

In k-fold cross validation, we first split up the data into $k$
folds,$X_1, X_2, ... , X_k$, each of length approximately $n/k$.

-   We then use the first fold $X_1$ of length $n/k$ as the test set,
    and the remaining $k-1$ folds combined, i.e. $X_2 + X_3 +... + X_k$
    of length \$ n - n/k\$, are used as the training set. The error rate
    is calculated, say $e_1$.

-   Then, the second fold fold $X_2$ is used as as the test set, and the
    remaining $k-1$ folds combined, i.e. $X_1 + X_3 +... + X_k$, are
    used as the training set. The error rate is calculated, say $e_2$

-   This step is repeated for all $k$ folds.

-   The average error rate is then used as the k-fold cross validation
    error rate: $$ e = \frac{1}{k} \sum_{i=1}^{k} e_i $$

### (b)

**What are the advantages and disadvantages of k-fold cross validation
relative to:**

#### **i. compared to the validation set approach?**

**Advantages of k-fold cross validation**

-   The error rate is less variable. In validation set approach, the
    error rate is highly variable and depends on the specific validation
    and testing set chosen.

-   When $k>2$, the k-fold cross validation yields more accurate
    estimates of the test error rate. This is because the validation set
    approach breaks up the data set into a validation set and a testing
    set. Due to lower number of observations in the validation set, the
    model gets to be trained on fewer observations. So, it overestimates
    the true error rates. In contrast, k-fold cross validation uses a
    larger training set, and somewhat ameliorates this situation.\

**Disadvantages of k-fold cross validation**

It is computationally more intensive, as the model needs to be fitted
$k$ times. But this is not a major concern in today's computers.

#### **ii. compared to LOOCV?**

**Advantages of k-fold cross validation**

-   As compared to LOOCV, k-fold cross validation is computationally
    less expensive (when $k < n$) because the model needs to be fitted a
    fewer number of times. A sole exception is linear regression, where
    a shortcut formula exists for calculating the LOOCV error rate using
    leverage value $h_i$ of each observation.

-   LOOCV averages the error rates over $n$ different models, each run
    on a training set consisting of $n-1$ observations. Since the
    training set of each LOOCV model is nearly identical, the error
    rates tend to be highly correlated. The average of highly correlated
    terms exhibits high variance. Thus, the LOOCV estimate suffers from
    high variance.

**Disadvantages of k-fold cross validation**

-   Each LOOCV model uses $n-1$ observations to fit the model, and thus
    uses a training set which is nearly identical to the original data
    set. Thus, LOOCV suffers from a very low bias. Hence, in terms of
    accuracy of the estimated error rate, LOOCV performs better when
    bias of the estimate is a major concern.

-   Overall, empirical evidence suggests that k-fold cross validation
    with $k=5$ or $k=10$ performs best considering this bias-variance
    trade off.

## Question 4

**Suppose that we use some statistical learning method to make a
prediction for the response Y for a particular value of the predictor X.
Carefully describe how we might estimate the standard deviation of our
prediction.**

To estimate the standard deviation of response $Y$ for a particular
value of predictor (say $X = x$), we can use the bootstrap method as
follows:

First, for some large number of repetitions $B$, we repeatedly take a
random sample with replacement from the observations of $X$. Lets call
these samples $X^{*1}, X^{*2} ... X^{*B}$. - We use the random samples
$X^{*1}, X^{*2} ... X^{*B}$ and their corresponding response values
($Y$) to train the statistical learning method one after the other. -
For each repetition we estimate the predicted value of response i.e.
$\hat{y^{*i}}$ for the particular value of predictor $X=x$. We term
these predictions as
$\hat{Y} = \hat{y^{*1}}, \hat{y^{*2}} ... \hat{y^{*B}}$.

Then we find the standard deviation of this estimated $\hat{Y}$. This
standard deviation is the bootstrap estimate for the true standard
deviation of response $Y$.
$$ SE_B(Y) = \sqrt {\frac{1}{B-1} \sum_{i=1}^{B} \big( \hat{y}^{*i} - \overline{\hat{y}}  \big)^2   }$$
where, $$ \overline{\hat{y}} = \frac{1}{B} \sum_{i'=1}^{B} y^{*i'}$$.\
This formula can be re-written as equation 5.8 from the text book:
$$ SE_B(Y) = \sqrt {\frac{1}{B-1} \sum_{i=1}^{B} \Big( \hat{y}^{*i} - \frac{1}{B} \sum_{i'=1}^{B} y^{*i'}  \Big)^2   }$$

# Applied

## Question 6

**In Chapter 4, we used logistic regression to predict the probability
of default using income and balance on the Default data set. We will now
estimate the test error of this logistic regression model using the
validation set approach. Do not forget to set a random seed before
beginning your analysis.**

### (a)

**Fit a logistic regression model that uses income and balance to
predict default.**

We fit the logistic regression model as shown below:

```{r}
library(ISLR)
library(kableExtra)
data("Default")
fit.glm <- glm(default ~ income + balance, data = Default, family = binomial)

# To print table
summary(fit.glm)$coef |>
  kbl(digits = 5) |>
  kableExtra::kable_material()
```

### (b)

**Using the validation set approach, estimate the test error of this
model. In order to do this, you must perform the following steps:**

We use a random vector `train` to divide equal number of observations
into the training set and validation set. We also `set seed` for
replicability. The results show that the validation set error rate is
2.64 %.

```{r}
set.seed(3)
# i. Split the sample set into a training set and a validation set.
train <- sample(x = 1:nrow(Default), size = nrow(Default) / 2)

# ii. Fit a multiple logistic regression model using only the training observations
fit.glm <- glm(default ~ income + balance,
  data = Default,
  family = binomial, subset = train
)

# Check whether 1 stands for Yes or No in default
contrasts(Default$default)

# iii. Obtain a prediction of default status for each individual
prob.glm <- predict(fit.glm,
  newdata = Default[-train, ],
  type = "response"
)
# Compute predicted response and create Confusion Matrix
pred.glm <- ifelse(prob.glm > 0.5, yes = "Yes", no = "No")
table(pred.glm, Default$default[-train])

# iv. Compute the validation set error rate
mean(pred.glm != Default$default[-train])
```

### (c)

**Repeat the process in (b) three times, using three different splits of
the observations into a training set and a validation set. Comment on
the results obtained.**

The 3 replications are done in the code below. The error rates obtained
for the validation set are 2.8 %, 2.74 % and 2.4 % which shows a fair
amount of variability in the results depending on which specific set of
observations are used for training and validation set.

```{r}
set.seed(3)
Error.Rates <- rep(NA, 3)
for (i in 1:3) {
  train <- sample(x = 1:nrow(Default), size = nrow(Default) / 2)
  fit.glm <- glm(default ~ income + balance,
    data = Default,
    family = binomial, subset = train
  )
  prob.glm <- predict(fit.glm,
    newdata = Default[-train, ],
    type = "response"
  )
  pred.glm <- ifelse(prob.glm > 0.5, yes = "Yes", no = "No")
  Error.Rates[i] <- mean(pred.glm != Default$default[-train])
}
Error.Rates
```

### (d)

**Now consider a logistic regression model that predicts the probability
of "default" using "income", "balance", and a dummy variable for
"student". Estimate the test error for this model using the validation
set approach. Comment on whether or not including a dummy variable for
"student" leads to a reduction in the test error rate.**

The new logistic regression model is fit as shown below, and the test
error using validation set approach is computed. The validation set
error rate is 2.5 %. This error rate is nearly same as before. Thus,
adding `student` variable in the logistic regression model does not lead
to any significant reduction in test error rate.

```{r}
# Checking which values represent student status in `student`
contrasts(Default$student)

# fitting multiple logistic regression model and computing predicted responses
fit.glm <- glm(default ~ income + balance + student,
  data = Default,
  family = binomial, subset = train
)
prob.glm <- predict(fit.glm,
  newdata = Default[-train, ],
  type = "response"
)
pred.glm <- ifelse(prob.glm > 0.5, yes = "Yes", no = "No")

# Display Confusion Matrix
table(pred.glm, Default$default[-train])

# Computing validation set error rate
mean(pred.glm != Default$default[-train])
```

## Question 6

**We continue to consider the use of a logistic regression model to
predict the probability of default using income and balance on the
Default data set. In particular, we will now compute estimates for the
standard errors of the income and balance logistic regression
coefficients in two different ways: (1) using the bootstrap, and (2)
using the standard formula for computing the standard errors in the
`glm()` function. Do not forget to set a random seed before beginning
your analysis.**

### (a)

**Using the `summary()` and `glm()` functions, determine the estimated
standard errors for the coefficients associated with income and balance
in a multiple logistic regression model that uses both predictors.**

The coefficients and their associated standard errors for `income` and
`balance` are shown below.

```{r, collapse=TRUE}
fit.glm <- glm(default ~ income + balance, 
               data = Default, 
               family = binomial)

# Displaying coefficients and standard errors
summary(fit.glm)$coef[2:3,1:2]
```

### (b)

**Write a function, `boot.fn()`, that takes as input the Default data
set as well as an index of the observations, and that outputs the
coefficient estimates for income and balance in the multiple logistic
regression model.**

We write the function in `R` code as below:---

```{r}
boot.fn <- function(data, index){
fit <- glm(default ~ income + balance, data = data, 
           family = binomial, subset = index)
return((fit)$coef)
}
# boot.fn(Default, 1:nrow(Default)) 
# Trial code to test the correctness of function
```

### (c)

**Use the `boot()` function together with your `boot.fn()` function to
estimate the standard errors of the logistic regression coefficients for
income and balance.**

We use the `boot` function from `boot` library in `R`to compute the
standard errors as follows:

```{r, cache=TRUE}
library(boot)
b1 = boot(data = Default, statistic = boot.fn, R = 1000)
b1
```

### (d)

**Comment on the estimated standard errors obtained using the glm()
function and using your bootstrap function.**

The coefficient estimates from the bootstrap and `glm()` function are
nearly identical. Similarly, the standard error for `income` coefficient
is 4.98e-6 and 5.02e-6, which are nearly the same. Lastly, the standard
error for `balance` coefficient is 2.27e-4 and 2.31e-4, which are nearly
same. Thus, it can be concluded that `glm()` and bootstrap are in
agreement. The logistic regression model fits the data quite well.

## Question 7

**In sections 5.3.2 and 5.3.3, we saw that the `cv.glm()` function can
be used in order to compute the LOOCV test error estimate.
Alternatively, one could compute those quantities using just the `glm()`
and `predict.glm()` functions, and a for loop. You will now take this
approach in order to compute the LOOCV error for a simple logistic
regression model on the "Weekly" data set. Recall that in the context of
classification problems, the LOOCV error is given in (5.4).**

### (a)

**Fit a logistic regression model that predicts "Direction" using "Lag1"
and "Lag2".**

```{r}
library(ISLR)
data(Weekly)
fit.glm <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = binomial)
```

### (b)

**Fit a logistic regression model that predicts Direction using Lag1 and
Lag2 using all but the first observation.**

```{r}
test <- 1
fit.glm <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = binomial,
               subset = -test)
```

### (c)

**Use the model from (b) to predict the direction of the first
observation. You can do this by predicting that the first observation
will go up if** $P(Direction=Up|Lag1, Lag2) > 0.5$. **Was this
observation correctly classified?**

The code given below predicts the `Direction` for the first observation.
It also shows that the first observation was incorrectly classified.

```{r, collapse=TRUE}
prob.glm <- predict(fit.glm, newdata = Weekly[test,], type = "response")

# Check the value of Up in dataset
contrasts(Weekly$Direction)
# Predict the response for observation no. 1
pred.glm <- ifelse(test = prob.glm > 0.5, yes = "Up", no = "Down")
pred.glm

# This observation was correctly classified: TRUE or FALSE
pred.glm == Weekly$Direction[test]
```

### (d)

**Write a for loop from i = 1 to i = n, where n is the number of
observations in the data set, that performs each of the following
steps**

The code is written below using `for` loops. The result is stored in the
vector `Errors`. The final result indicates that nearly 45% of the
looped models resulted in an erroneous prediction.

```{r}
# Create a vector Errors to store results of loops
Errors <- rep(NA, nrow(Weekly))
# Creating for loop from 1 to n
for (i in 1:nrow(Weekly)) {
  test <- i
  
  # i. Fit a logistic regression model using all but the ith observation 
  fit.glm <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = binomial,
               subset = -test)
  # ii. Compute the posterior probability of the market moving "Up" for 
  #     the ith observation.
  prob.glm <- predict(fit.glm, newdata = Weekly[test,], type = "response")
  
  # iii. Use the posterior probability for the ith observation in order 
  #      to predict whether or not the market moves up.
  pred.glm <- ifelse(test = prob.glm > 0.5, yes = "Up", no = "Down")
  
  # iv. Determine whether or not an error was made in predicting
  #     the direction for the ith observation. If an error was made,
  #     then indicate this as a 1, and otherwise indicate it as a 0.
  Errors[i] <- ifelse(pred.glm != Weekly$Direction[test], yes = 1, no = 0)
}
mean(Errors)
```

### (e)

**Take the average of the n numbers obtained in (d) in order to obtain
the LOOCV estimate for the test error. Comment on the results.**

As shown above. the `mean(Errors)` shows us LOOCV estimate for test
error rate as 44.995 %. This is an exceptionally high error rate, and
only slightly better than random guessing error rate of 50 %.

## Question 8

**We will now perform cross-validation on a simulated data set.**

### (a)

**In this data set, what is n and what is p? Write out the model used to
generate the data in equation form.**

We generate a simulated data set as follows, based on the code given in
the question. In this data set, $n = 100$ and $p=2$. The model used to
generate the data in equation form is:
$$ y = x \ - \ 2x^2 \ + \ \epsilon$$

```{r}
set.seed(1)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
```

### (b)

**Create a scatter-plot of X against Y . Comment on what you find.**

The scatter-plot of $y$ vs. $x$ is shown below. It shows a prominent
non-linear pattern between $x$ and $y$.

```{r}
plot(x,y)
```

### (c)

**Set a random seed, and then compute the LOOCV errors that result from
fitting the following four models using least squares:**

The code below uses `for` loops to calculate LOOCV error rate for the
fitted models involving polynomials of $x$. The results are displayed in
the `Result` data-frame at the end. It is clear the the LOOCV error rate
of model (i) is very high, and thereafter the other models have similar
error rates. There seems no benefit of adding $x^3$ and $x^4$ to the
model.

```{r}
library(boot)
# Setting random seed
set.seed(3)

# Creating Data Frames for x and y, and for storing final result
Dataset <- data.frame(x, y)
Result <- data.frame(
  Model = c("i", "ii", "iii", "iv"),
  ErrorRate = rep(NA, 4)
)

# Using loops to calculate LOOCV Error Rate
for (i in 1:4) {
  fit <- glm(y ~ poly(x, i), data = Dataset)
  Result[i, 2] <- cv.glm(data = Dataset, fit)$delta[1]
}

# Displaying the error rates for all four models
Result
```

### (d)

**Repeat (c) using another random seed, and report your results. Are
your results the same as what you got in (c)? Why?**

We now repeat the results using another seed, lets say `set.seed(9)`.
The results are computed below. The resulting error rates are exactly
the same irrespective of the random seed. This is expected because LOOCV
eliminates the aspect of randomness by using every single observation as
a validation set. Thus, LOOCV estimate will never vary based on random
seed. Further, in case of linear regression, LOOCV estimates can be
computed with a short cut formula which does not depend on random
sampling at all.

```{r}
library(boot)
set.seed(9)
Dataset <- data.frame(x, y)
Result1 <- data.frame(
  Model = c("i", "ii", "iii", "iv"),
  ErrorRate = rep(NA, 4)
)
for (i in 1:4) {
  fit <- glm(y ~ poly(x, i), data = Dataset)
  Result1[i, 2] <- cv.glm(data = Dataset, fit)$delta[1]
}
Result1 |> kbl()
```

### (e)

**Which of the models in (c) had the smallest LOOCV error? Is this what
you expected? Explain your answer.**

In the models computed in (c) above, the model with $x$ and $x^2$, i.e.
model (ii) has the lowest LOOCV error. This is expected, as the true
model is same as the estimate model (ii). We had created the true model
in part (a).

### (f)

**Comment on the statistical significance of the coefficient estimates
that results from fitting each of the models in (c) using least squares.
Do these results agree with the conclusions drawn based on the
cross-validation results?**

We can compute the coefficient estimates from fitting the models again
as follows, and print the `summary(fit)` to examine the statistical
significance of each predictor. The predictors $x$ and $x^2$ are
statistically significant predictors in all models. However, the
predictors $x^3$ and $x^4$ are not statistically significant. These
results agree with the conclusions drawn based on cross-validation
results which showed that the error rates for models using $x^3$ and
$x^4$ are not any better than previous models. Adding the terms $x^3$
and $x^4$ did not lower the error rate.

```{r}
for (i in 1:4) {
  fit <- lm(y ~ poly(x, i), data = Dataset)
  print(summary(fit))
}
```

## Question 9

### (a)

The estimate for population mean of `medv` is 22.53. This is the sample
mean $\hat{\mu}$ which is an unbiased estimate for $\mu$, the true
population mean. The `R` code to compute it is given below.

```{r}
library(MASS)
data(Boston)
attach(Boston)
mean(medv)
```

### (b)

The estimated standard error for $\hat{\mu}$ can be given by
$sd/\sqrt{n}$, where $sd$ stands for standard deviation of the random
variable whose true mean is $\mu$ and $n$ represents number of
observations in the sample. The `R` code below calculates this estimated
standard error as 0.409.

```{r}
sd(medv)/(sqrt(nrow(Boston)))
```

### (c)

We now compute the standard error of estimated mean of `medv` using
bootstrap. The estimated standard error using bootstrap method is 0.408,
which is reasonably close to the previously calculated standard error of
0.409.

```{r, collapse=TRUE}
library(boot)

# Creating a function which returns mean of medv as output
boot.fn <- function(data, index) {
  return(mean(data$medv[index]))
}

# Testing the boot.fn on whole Boston data set to ensure it works
boot.fn(Boston, 1:nrow(Boston))

# Using bootstrap to calculate standard error
set.seed(3)
boot(data = Boston, statistic = boot.fn, R = 10000)

```

### (d)

We compute the 95% confidence interval for mean of `medv` using
bootstrap standard error estimates. The resulting 95% Confidence
Interval is \[21.72 , 23.35\]. The 95 % confidence interval computed
using the `t.test` function in `R` is \[21.73 , 23.34\]. Thus, we
conclude the the two confidence intervals are nearly identical.

```{r}
mean(medv) - 2 *0.408
mean(medv) + 2 *0.408

# Using t.test to calculate the 95 % Confidence Interval
t.test(Boston$medv)
```

### (e)

The estimated median value of `medv` in the population is the median of
the `Boston` sample data set. This $\hat{\mu}$ value is 21.2 as
calculated below.

```{r}
median(Boston$medv)
```

### (f)

As stated in the question, there is no easy formula to compute the
standard error associated with a estimated sample median. Instead, we
can use bootstrap method to compute the estimated standard error
associated with the sample median. The estimated standard error is 0.37.

```{r}
library(boot)

# Creating a function which returns median of medv as output
boot.fn1 <- function(data, index){
  return(median(data$medv[index]))
}

# Testing the boot.fn on whole Boston data set to ensure it works
boot.fn1(Boston, 1:nrow(Boston))

# Using bootstrap to calculate standard error
set.seed(3)
boot(data = Boston, statistic = boot.fn1, R = 10000)
```

### (g)

We can compute the estimate for 10\^{th} percentile of true population
`medv` using the `quantile()` function of the `medv` in `Boston` data
set. The estimated $\hat{\mu_{0.1}}$ is 12.75 as shown below.

```{r}
quantile(Boston$medv, 0.1)
```

### (h)

Again, there is no easy mathematical formula to calculate estimated
standard error associated with 10^th^ percentile of a sample. We can use
bootstrap to calculate the estimated standard error as shown below. The
estimated standard error for $\hat{\mu_{0.1}}$ is 0.5.

```{r}
library(boot)

# Creating a function which returns 10th percentile of medv as output
boot.fn2 <- function(data, index) {
  return(quantile(data$medv[index], 0.1))
}

# Testing the boot.fn on whole Boston data set to ensure it works
boot.fn2(Boston, 1:nrow(Boston))

# Using bootstrap to calculate standard error
set.seed(3)
boot(data = Boston, statistic = boot.fn2, R = 10000)
```
